Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-11-08 Manuscript/manuscript.tex
Small edits to the abstract

\begin{abstract}
  % insert abstract here
Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
bilayers are of paramount significance in the properties and functionality of cellular
membranes. For example, possible accumulation of ions at the membrane interface determines
the interaction affinity with charged species. Current estimates of binding affinities and
stoichiometries of cations are controversial due to limitations in the experimental and
theoretical methods available. Experimentally one can assess these parameters by titrating
several membrane properties as a function of the ions concentration. However, the
interpretation of the obtained results requires theoretical models. Considering that the
real behavior is unknown these models are at best idealizations. Classical molecular
dynamics (MD) simulations provide details of the ions binding process with atomistic
resolution. Therefore, MD a priory offers all necessary information to interpret
experimental data without the need to resort to simplified models. However, the accuracy
of the available lipid models when interacting with ions is currently limited. In this
work, we improve the binding details of Na$^+$ and Ca$^{2+}$ ions to the
1-Palmitoyl-2-oleoyl-phosphatidylcholine POPC) bilayer by implicitly including electronic
polarization. We do this by applying a mean field correction to the partial charges known
as the electronic continuum correction (ECC) to a selected state of the art POPC lipid
model for MD simulations. Our improved ECC-POPC model reproduces not only the
experimentally measured structural parameters for the ion-free membrane, but also the
response of lipid headgroup to a bound positive charge, and the binding affinities of 
Na$^+$ and Ca$^{2+}$ ions. We also observe negligible binding of Na$^+$ ions to POPC
bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens in our new
model is in agreement with available experimental spectroscopic data. This new model
results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
equal probabilities which suggests that previous models used to interpret NMR data are too
simple. The results of this work pave the way to quantitative MD simulations of complex
biochemical systems with realistic electrostatic interactions in the vicinity of cellular
membranes.
\end{abstract}

>>>>

\begin{abstract}
  % insert abstract here
Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
bilayers are of paramount significance in the properties and functionality of cellular
membranes. For example, possible accumulation of ions at the membrane interface determines
the interaction affinity with charged species. Current estimates of binding affinities and
stoichiometries of cations are, however, controversial due to limitations in the available experimental and
theoretical methods. Experimentally one can assess these parameters by titrating
several membrane properties as a function of the ion concentration. However, the
interpretation of the experiments relies on theoretical models, which give
the best idealization of the ion binding as the direct experiments are not available. 
Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic
resolution, therefore offering all the necessary information to interpret
experimental data without the need to resort to simplified models. However, the accuracy
of the available lipid models when interacting with ions is not sufficient for
such interpretation. In this work, we improve the binding details of Na$^+$ and Ca$^{2+}$ ions to the
1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
polarization as a mean field correction, known as the electronic continuum correction (ECC).
This is applied to the partial charges of a selected state of the art POPC lipid model for MD simulations.
Our improved ECC-POPC model reproduces not only the
experimentally measured structural parameters for the ion-free membrane, but also the
response of lipid headgroup to a bound positive charge, and the binding affinities of 
Na$^+$ and Ca$^{2+}$ ions. With our new model we also observe negligible binding of Na$^+$ ions to POPC
bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens 
, which are in agreement with the previous interpretations of the experimental spectroscopic data.
On the other hand, the new model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
equal probabilities, suggesting more complex binding stoichiometry than simple models
previously used to interpret the NMR data.
The results of this work pave the way to quantitative MD simulations of complex
biochemical systems with realistic electrostatic interactions in the vicinity of cellular
membranes.
\end{abstract}
